---
title: "Checking Out Data"
author: "Theo"
date: "10/16/2024"
output: html_document
---








#DATA VISUALIZATION CODE BITS
#loading libraries
``{r}
library(tidyverse)
library(ggplot2)
library(scales)
``




#readfiles
``{r}
data <- read_csv("your_data.csv")
``



#basic scatterplot (labs = labels)

``{r}
ggplot(data, aes(x = x_variable, y = y_variable)) +
  geom_point() +
  labs(title = "Scatter Plot", x = "X-axis Label", y = "Y-axis Label")
``



#line plots

``{r}
ggplot(data, aes(x = x_variable, y = y_variable)) +
  geom_line() +
  labs(title = "Line Plot", x = "X-axis Label", y = "Y-axis Label")

``


#histogram

``{r}
ggplot(data, aes(x = numeric_variable)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Histogram", x = "Values", y = "Frequency")
``


#boxplot
``{r}
ggplot(data, aes(x = category_variable, y = numeric_variable)) +
  geom_boxplot() +
  labs(title = "Box Plot", x = "Categories", y = "Values")
``


#add color to plots

``{r}
ggplot(data, aes(x = x_variable, y = y_variable, color = group_variable)) +
  geom_point() +
  labs(title = "Colored Scatter Plot", x = "X-axis Label", y = "Y-axis Label")
``


# CODE FOR CSV FILES






#view structure of data
``{r}
str(data)
summary(data)
``


#select specific columns

``{r}
library(dplyr)
selected_data <- data %>% select(column1, column2, column3)
``


#Filter rows based on condition
``{r}
filtered_data <- data %>% filter(column1 > 10)
``



#create new column
``{r}
data <- data %>% mutate(new_column = column1 + column2)
``

#rename columns
``{r}
data <- data %>% rename(new_name = old_name)
``


#group data and make summary stats
``{r}
summary_data <- data %>%
  group_by(category_column) %>%
  summarise(
    mean_value = mean(numeric_column),
    max_value = max(numeric_column)
  )
``






```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<<<<<<< HEAD
=======
# seeing whats up with our data and how big it is, also if its any good
>>>>>>> bf4bfc1298d105fd0e414269a4b9ee1e950bd231

```{r}
library(readr)
```


#changing names and deleting belgium-luxembourg 
```{r}

afgan_cambo <- read_csv("RAW_CSV_DATA/afgan-cambo.csv")

afgan_cambo$Area[afgan_cambo$Area == "Bolivia (Plurinational State of)"] <- "Bolivia"

afgan_cambo$Area[afgan_cambo$Area == "Brunei Darussalam"] <- "Brunei"
 
afgan_cambo$Area[afgan_cambo$Area == "Cabo Verde"] <- "Cape Verde"

afgan_cambo <- afgan_cambo[afgan_cambo$Area != "Belgium-Luxembourg", ]

afgan_cambo <- read.csv("RAW_CSV_DATA/afgan-cambo.csv")

write.csv(afgan_cambo, file = "fully_afgan_cambo_cleaned.csv")

```

```{r}
camer_elsalv <- read_csv("RAW_CSV_DATA/camer-elsalv.csv")

camer_elsalv <- camer_elsalv[camer_elsalv$Area != "China, mainland", ]

camer_elsalv <- camer_elsalv[camer_elsalv$Area != "China, Macao SAR", ]

camer_elsalv$Area[camer_elsalv$Area == "China, Hong Kong SAR"] <- "Hong Kong"
 
camer_elsalv$Area[camer_elsalv$Area == "China, Taiwan Province of"] <- "Taiwan"

camer_elsalv$Area[camer_elsalv$Area == "Côte d'Ivoire"] <- "Ivory Coast"

camer_elsalv$Area[camer_elsalv$Area == "Czechia"] <- "Czech Republic"

camer_elsalv$Area[camer_elsalv$Area == "Democratic People's Republic of Korea"] <- "North Korea"

camer_elsalv <- read.csv("RAW_CSV_DATA/camer-elsalv.csv")

write.csv(camer_elsalv, file = "fully_cleaned_camer_elsalv.csv")

```





```{r}

Tanz_Zimb <- read_csv("RAW_CSV_DATA/Tanz-Zimb.csv")

Tanz_Zimb$Area[Tanz_Zimb$Area == "United Republic of Tanzania"] <- "Tanzania"

Tanz_Zimb$Area[Tanz_Zimb$Area == "Venezuela (Bolivarian Republic of)"] <- "Venezuela"

Tanz_Zimb$Area[Tanz_Zimb$Area == "Viet Nam"] <- "Vietnam"

Tanz_Zimb <- Tanz_Zimb[Tanz_Zimb$Area != "Yugoslav SFR", ]

Tanz_Zimb <- read.csv("RAW_CSV_DATA/Tanz-Zimb.csv")

write.csv(Tanz_Zimb, file = "fully_tanz_zimb_cleaned.csv")

```

#changed Tanz_Zimb names (Vietnam, Tanzania, Venezuela)
