---
title: "Checking Out Data"
author: "Theo"
date: "10/16/2024"
output: html_document
---








#DATA VISUALIZATION CODE BITS
#loading libraries
``{r}
library(tidyverse)
library(ggplot2)
library(scales)
``




#readfiles
``{r}
data <- read_csv("your_data.csv")
``



#basic scatterplot (labs = labels)

``{r}
ggplot(data, aes(x = x_variable, y = y_variable)) +
  geom_point() +
  labs(title = "Scatter Plot", x = "X-axis Label", y = "Y-axis Label")
``



#line plots

``{r}
ggplot(data, aes(x = x_variable, y = y_variable)) +
  geom_line() +
  labs(title = "Line Plot", x = "X-axis Label", y = "Y-axis Label")

``


#histogram

``{r}
ggplot(data, aes(x = numeric_variable)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Histogram", x = "Values", y = "Frequency")
``


#boxplot
``{r}
ggplot(data, aes(x = category_variable, y = numeric_variable)) +
  geom_boxplot() +
  labs(title = "Box Plot", x = "Categories", y = "Values")
``


#add color to plots

``{r}
ggplot(data, aes(x = x_variable, y = y_variable, color = group_variable)) +
  geom_point() +
  labs(title = "Colored Scatter Plot", x = "X-axis Label", y = "Y-axis Label")
``


# CODE FOR CSV FILES






#view structure of data
``{r}
str(data)
summary(data)
``


#select specific columns

``{r}
library(dplyr)
selected_data <- data %>% select(column1, column2, column3)
``


#Filter rows based on condition
``{r}
filtered_data <- data %>% filter(column1 > 10)
``



#create new column
``{r}
data <- data %>% mutate(new_column = column1 + column2)
``

#rename columns
``{r}
data <- data %>% rename(new_name = old_name)
``


#group data and make summary stats
``{r}
summary_data <- data %>%
  group_by(category_column) %>%
  summarise(
    mean_value = mean(numeric_column),
    max_value = max(numeric_column)
  )
``



# seeing whats up with our data and how big it is, also if its any good

```{r}
library(readr)
```


#changing names in afgan-cambo raw data
```{r}

afgan_cambo <- read.csv("afgan-cambo-raw.csv")

afgan_cambo$Area[afgan_cambo$Area == "Bolivia (Plurinational State of)"] <- "Bolivia"

afgan_cambo$Area[afgan_cambo$Area == "Brunei Darussalam"] <- "Brunei"
 
afgan_cambo$Area[afgan_cambo$Area == "Cabo Verde"] <- "Cape Verde"

afgan_cambo <- afgan_cambo[afgan_cambo$Area != "Belgium-Luxembourg", ]

write.csv(afgan_cambo, file = "fully_afgan_cambo_cleanedV2.csv")

```



#changing names in camer-elsalv raw data (Check Czecholslovakia)
```{r}
camer_elsalv <- read.csv("RAW_CSV_DATA/camer-elsalv.csv")

camer_elsalv <- camer_elsalv[camer_elsalv$Area != "China, mainland", ]

camer_elsalv <- camer_elsalv[camer_elsalv$Area != "China, Macao SAR", ]

camer_elsalv$Area[camer_elsalv$Area == "China, Hong Kong SAR"] <- "Hong Kong"
 
camer_elsalv$Area[camer_elsalv$Area == "China, Taiwan Province of"] <- "Taiwan"

camer_elsalv$Area[camer_elsalv$Area == "Côte d'Ivoire"] <- "Ivory Coast"

camer_elsalv$Area[camer_elsalv$Area == "Czechia"] <- "Czech Republic"

camer_elsalv$Area[camer_elsalv$Area == "Democratic People's Republic of Korea"] <- "North Korea"

write.csv(camer_elsalv, file = "fully_camer_elsalv_cleaned.csv")

```


#changing names in eqgu-hung raw data
```{r}
eqgu_hung <- read.csv("RAW_CSV_DATA/eqgu-hung.csv")

eqgu_hung$Area[eqgu_hung$Area == "Ethiopia PDR"] <- "Ethiopia"

write.csv(eqgu_hung, file = "fully_eqgu_hung_cleaned.csv")

```


#changing names in ice-lib raw data
```{r}
ice_lib <- read.csv("RAW_CSV_DATA/ice-lib.csv")

ice_lib$Area[ice_lib$Area == "Iran (Islamic Republic of)"] <- "Iran"

ice_lib$Area[ice_lib$Area == "Lao People's Democratic Republic"] <- "Laos"

write.csv(ice_lib, file = "fully_ice_lib_cleaned.csv")

```


#changing names in lith-monte raw data
```{r}
lith_monte <- read.csv("RAW_CSV_DATA/lith-monte.csv")

lith_monte$Area[lith_monte$Area == "Micronesia (Federated States of)"] <- "Micronesia"

write.csv(lith_monte, file = "fully_lith_monte_cleaned.csv")

```


#changing names in moro-roma raw data
```{r}
moro_roma <- read.csv("RAW_CSV_DATA/moro-roma.csv")

moro_roma$Area[moro_roma$Area == "Netherlands (Kingdom of the)"] <- "Netherlands"

moro_roma$Area[moro_roma$Area == "Republic of Korea"] <- "South Korea"

moro_roma$Area[moro_roma$Area == "Republic of Moldova"] <- "Moldova"

write.csv(moro_roma, file = "fully_moro_roma_cleaned.csv")

```


#changing names in russfed-swed raw data  (need to change sudan (former) and serbia/montenegro)
```{r}
russfed_swed <- read.csv("RAW_CSV_DATA/russfed-swed.csv")

russfed_swed$Area[russfed_swed$Area == "Russian Federation"] <- "Russia"

write.csv(russfed_swed, file = "fully_russfed_swed_cleaned.csv")

```


#changing names in Switz-UK raw data
```{r}
switz_uk <- read.csv("RAW_CSV_DATA/Switz-UK.csv")

switz_uk$Area[switz_uk$Area == "Syrian Arab Republic"] <- "Syria"

switz_uk$Area[switz_uk$Area == "Timor-Leste"] <- "East Timor"

switz_uk$Area[switz_uk$Area == "Türkiye"] <- "Turkey"

switz_uk$Area[switz_uk$Area == "United Kingdom of Great Britain and Northern Ireland"] <- "United Kingdom"

write.csv(switz_uk, file = "fully_switz_uk_cleaned.csv")

```


#changing names in Tanz-Zimb raw data 
```{r}

Tanz_Zimb <- read.csv("RAW_CSV_DATA/Tanz-Zimb.csv")

Tanz_Zimb$Area[Tanz_Zimb$Area == "United Republic of Tanzania"] <- "Tanzania"

Tanz_Zimb$Area[Tanz_Zimb$Area == "Venezuela (Bolivarian Republic of)"] <- "Venezuela"

Tanz_Zimb$Area[Tanz_Zimb$Area == "Viet Nam"] <- "Vietnam"

Tanz_Zimb <- Tanz_Zimb[Tanz_Zimb$Area != "Yugoslav SFR", ]

write.csv(Tanz_Zimb, file = "fully_tanz_zimb_cleaned.csv")

```

#changed Tanz_Zimb names (Vietnam, Tanzania, Venezuela)






#changing names in fully_temp_data_clean from raw Temperature-Data to match names in cleaned agriculture data
```{r}
temp_data <- read.csv("RAW_CSV_DATA/Temperature-Data.csv")

temp_data$Country[temp_data$Country == "Afghanistan, Islamic Rep. of"] <- "Afghanistan"

temp_data <- temp_data[temp_data$Country != "American Samoa", ]

temp_data <- temp_data[temp_data$Country != "Palau, Rep. of", ]

temp_data <- temp_data[temp_data$Country != "Falkland Islands (Malvinas)", ]

temp_data <- temp_data[temp_data$Country != "Andorra, Principality of", ]

temp_data <- temp_data[temp_data$Country != "Anguilla", ]

temp_data$Country[temp_data$Country == "Armenia, Rep. of"] <- "Armenia"

temp_data$Country[temp_data$Country == "Taiwan Province of China"] <- "Taiwan"

temp_data <- temp_data[temp_data$Country != "Aruba, Kingdom of the Netherlands", ]

temp_data$Country[temp_data$Country == "Azerbaijan, Rep. of"] <- "Azerbaijan"

temp_data$Country[temp_data$Country == "Bahrain, Kingdom of"] <- "Bahrain"

temp_data <- temp_data[temp_data$Country != "Bahamas, The", ]

temp_data$Country[temp_data$Country == "Belarus, Rep. of"] <- "Belarus"

temp_data <- temp_data[temp_data$Country != "British Virgin Islands", ]

temp_data$Country[temp_data$Country == "Brunei Darussalam"] <- "Brunei"

temp_data$Country[temp_data$Country == "Cabo Verde"] <- "Cape Verde"

temp_data$Country[temp_data$Country == "Central African Rep."] <- "Central African Republic"

temp_data$Country[temp_data$Country == "China, P.R.: Hong Kong"] <- "Hong Kong"

temp_data$Country[temp_data$Country == "China, P.R.: Mainland"] <- "China"

temp_data$Country[temp_data$Country == "Comoros, Union of the"] <- "Comoros"

temp_data$Country[temp_data$Country == "Congo, Rep. of"] <- "Congo"

temp_data$Country[temp_data$Country == "Congo, Dem. Rep. of the"] <- "Democratic Republic of the Congo"

temp_data$Country[temp_data$Country == "Côte d'Ivoire"] <- "Ivory Coast"

temp_data$Country[temp_data$Country == "Croatia, Rep. of"] <- "Croatia"

temp_data$Country[temp_data$Country == "Czech Rep."] <- "Czech Republic"

temp_data$Country[temp_data$Country == "Dominican Rep."] <- "Dominican Republic"

temp_data$Country[temp_data$Country == "Egypt, Arab Rep. of"] <- "Egypt"

temp_data$Country[temp_data$Country == "Equatorial Guinea, Rep. of"] <- "Equatorial Guinea"

temp_data$Country[temp_data$Country == "Eritrea, The State of"] <- "Eritrea"

temp_data$Country[temp_data$Country == "Estonia, Rep. of"] <- "Estonia"

temp_data$Country[temp_data$Country == "Eswatini, Kingdom of"] <- "Eswatini"

temp_data$Country[temp_data$Country == "Ethiopia, The Federal Dem. Rep. of"] <- "Ethiopia"

temp_data$Country[temp_data$Country == "St. Kitts and Nevis"] <- "Saint Kitts and Nevis"

temp_data$Country[temp_data$Country == "Fiji, Rep. of"] <- "Fiji"

temp_data$Country[temp_data$Country == "Guiana, French"] <- "French Guiana"

temp_data$Country[temp_data$Country == "Gambia, The"] <- "Gambia"

temp_data$Country[temp_data$Country == "Iran, Islamic Rep. of"] <- "Iran"

temp_data$Country[temp_data$Country == "Kazakhstan, Rep. of"] <- "Kazakhstan"

temp_data$Country[temp_data$Country == "South Sudan, Rep. of"] <- "South Sudan"

temp_data$Country[temp_data$Country == "St. Vincent and the Grenadines"] <- "Saint Vincent and the Grenadines"

temp_data$Country[temp_data$Country == "Slovenia, Rep. of"] <- "Slovenia"

temp_data$Country[temp_data$Country == "Slovak Rep."] <- "Slovakia"

temp_data$Country[temp_data$Country == "Serbia, Rep. of"] <- "Serbia"

temp_data$Country[temp_data$Country == "Korea, Dem. People's Rep. of"] <- "North Korea"

temp_data$Country[temp_data$Country == "Korea, Rep. of"] <- "South Korea"

temp_data$Country[temp_data$Country == "Kyrgyz Rep."] <- "Kyrgyzstan"

temp_data$Country[temp_data$Country == "Lao People's Dem. Rep."] <- "Laos"

temp_data$Country[temp_data$Country == "Lesotho, Kingdom of"] <- "Lesotho"

temp_data$Country[temp_data$Country == "São Tomé and Príncipe, Dem. Rep. of"] <- "Sao Tome and Principe"

temp_data$Country[temp_data$Country == "St. Lucia"] <- "Saint Lucia"

temp_data$Country[temp_data$Country == "Russian Federation"] <- "Russia"

temp_data$Country[temp_data$Country == "Türkiye, Rep. of"] <- "Turkey"

temp_data$Country[temp_data$Country == "Timor-Leste, Dem. Rep. of"] <- "East Timor"

temp_data$Country[temp_data$Country == "Tajikistan, Rep. of"] <- "Tajikistan"

temp_data$Country[temp_data$Country == "Syrian Arab Rep."] <- "Syria"

temp_data$Country[temp_data$Country == "Yemen, Rep. of"] <- "Yemen"

temp_data$Country[temp_data$Country == "Venezuela, Rep. Bolivariana de"] <- "Venezuela"

temp_data$Country[temp_data$Country == "Uzbekistan, Rep. of"] <- "Uzbekistan"

temp_data$Country[temp_data$Country == "United States"] <- "United States of America"

temp_data$Country[temp_data$Country == "Tanzania, United Rep. of"] <- "Tanzania"

temp_data$Country[temp_data$Country == "Madagascar, Rep. of"] <- "Madagascar"

temp_data$Country[temp_data$Country == "Marshall Islands, Rep. of the"] <- "Marshall Islands"

temp_data$Country[temp_data$Country == "Mauritania, Islamic Rep. of"] <- "Mauritania"

temp_data$Country[temp_data$Country == "Micronesia, Federated States of"] <- "Micronesia"

temp_data$Country[temp_data$Country == "Moldova, Rep. of"] <- "Moldova"

temp_data$Country[temp_data$Country == "Poland, Rep. of"] <- "Poland"

temp_data$Country[temp_data$Country == "West Bank and Gaza"] <- "Palestine"

temp_data$Country[temp_data$Country == "North Macedonia, Republic of "] <- "North Macedonia"

temp_data$Country[temp_data$Country == "Netherlands, The"] <- "Netherlands"

temp_data$Country[temp_data$Country == "Nauru, Rep. of"] <- "Nauru"

temp_data$Country[temp_data$Country == "Mozambique, Rep. of"] <- "Mozambique"

temp_data <- temp_data[temp_data$Country != "Cayman Islands", ]

temp_data <- temp_data[temp_data$Country != "China, P.R.: Macao", ]

temp_data <- temp_data[temp_data$Country != "Greenland", ]

temp_data <- temp_data[temp_data$Country != "Svalbard Jan Mayen Islands", ]

temp_data <- temp_data[temp_data$Country != "Holy See", ]

temp_data <- temp_data[temp_data$Country != "Isle of Man", ]

temp_data <- temp_data[temp_data$Country != "San Marino, Rep. of", ]

temp_data <- temp_data[temp_data$Country != "Saint Helena", ]

temp_data <- temp_data[temp_data$Country != "Saint Pierre and Miquelon", ]

temp_data <- temp_data[temp_data$Country != "Western Sahara", ]

temp_data <- temp_data[temp_data$Country != "Wallis and Futuna Islands", ]

temp_data <- temp_data[temp_data$Country != "United States Virgin Islands", ]

temp_data <- temp_data[temp_data$Country != "Mayotte", ]

temp_data <- temp_data[temp_data$Country != "Montserrat", ]

temp_data <- temp_data[temp_data$Country != "Pitcairn Islands", ]

temp_data <- temp_data[temp_data$Country != "Norfolk Island", ]

write.csv(temp_data, file = "fully_temp_data_cleaned.csv")

```